---
output: github_document
---

# Likert Survey Analysis


## Loading the libraries


```{r message=FALSE, warning=FALSE}
library(dplyr)
library(janitor)
library(tidyr)
library(ggplot2)
library(readxl)
library(stringr)
```


## Part 1

Loading the data that is in the `data` folder. 


```{r warning=FALSE}

students <- read_excel("data/likert_survey.xlsx", sheet = "students") %>%
  clean_names() 

professors <- read_excel("data/likert_survey.xlsx", sheet = "professors")%>%
  clean_names()

```



## Part 2

We are now looking into the total number of answers obtained by students, professors and if each question obtained the same number of answers.

To answer the first two questions I am using the `tidyverse` way, putting together the dataset, to then create long format data. This will allow me to exploit the power of grouping and summarising.

```{r}
# let's first put the two datasets together

students <- students %>% 
  mutate(respondent = "student") # create a respondent column

professors <- professors %>% 
  mutate(respondent = "professor") # create a respondent column

# We can now bind together the two tibbles

total <- bind_rows(students, professors)

# use head(total) to inspect the element

# And now we can use pivot_longer to get it in a tidy format

total <- total %>% 
  pivot_longer(names_to = "response_type",
               values_to = "value",
               cols= c(-question_id, -respondent))

head(total)
```

After having wrangled the data to have it in tidy format, I can respond to the first two questions.


```{r}
total %>% 
  group_by(respondent) %>% 
  summarise(total_answer = sum(value)) %>% 
  ungroup() %>% 
  rename('Total Answer' = total_answer,
         Respondent = respondent) %>% 
  knitr::kable()
```

We are now checking if each question have the same number of answers. 


```{r}
total %>% 
  group_by(respondent, question_id) %>%  # we group_by respondent and question_id
  summarise(total_answer = sum(value)) %>% # we calculate the total answer per question and respondent
  ungroup() %>% # we ungroup!
  distinct(respondent, total_answer) # we keep only the unique values

```
The answer is *No* as we can see that not all questions have been answered the same amount of time, for both professors and students. It means that some respondents left some questions unanswered. 


## Part 3


I am now comparing the distribution of answers of the students versus the professors.



```{r}
total %>%
  ggplot(mapping = aes(x =response_type,
                       y = value,
                       fill = respondent)) +
  geom_col(position = "dodge") + 
  facet_wrap(vars(question_id)) + 
  labs(title = "Comparaison of answers for students vs professors",
       subtitle = "Each facet shows a different question",
       x="Response type", 
       y="Number of Responses",
       fill = "respondent") +
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 90)) # tilting the text

```

From this chart, we can see that basically for each question, students gave more answers.

## Part 4

Now I am computing the mode for the students and professors for each of the questions. 

```{r}
total %>%
  group_by(question_id, respondent) %>%
  slice_max(order_by = value, n = 1) %>% 
  ungroup()

```
In order to make the data nice to present in a table we can select `response_type`, `respondent` and `question_id` and use `pivot_wider()` so to have one line per `question_id`. 

```{r}
total %>%
  group_by(question_id, respondent) %>%
  slice_max(order_by = value, n = 1) %>% # we keep the top 1 value for each group
  ungroup() %>% 
  select(-value) %>% 
  pivot_wider(names_from = respondent, values_from = response_type)

```


